The problem and promise of scale in multilayer networks

### 2020-02-03 paper/manuscript.Rmd
minor edits and proper todos

# Abstract
Scale remains a seminal concept in ecology. Spatial scale, for instance, has
become a central consideration in the way we understand landscape ecology and
animal space use. Meanwhile, social processes can also scale from fine-scale
interactions to co-occurrence to overlapping home ranges. Multilayer networks
promise the integration of monolayer animal social networks with the complexity
and importance of animal space use and movement in heterogeneous landscapes.
Despite the complex interplay between social networks and how animals use space,
there remains an important biological and methodological gap in our
understanding of an animal’s perception of scale, e.g., grain and extent. But
scale can affect multilayer network dynamics and how we build and interpret
them. Here, we discuss the role of scale in the context of multilayer networks
and provide a case study of caribou (*Rangifer tarandus*) to illustrate the role
of social,spatial, and temporal scale on multilayer processes. We also present
perspectives on future development and application of scale in multilayer
networks with respect to habitat selection and space use, individual fitness and
phenotypes, and movement ecology and collective movement.Effective integration
of social and spatial processes, including biologically meaningful scales,
within the context of animal social networks is an emerging area of research;
our contribution uniquely incorporates perspectives on how the social
environment and spatial processes are linked across scales in a multilayer
framework. Based on social network and behavioural ecology theory as well as our
case study, we demonstrate potential for interconnectedness among individual
animals, their social environment, and the landscapes they occupy.

@@

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address two
principal questions associated with scale-dependent multilayer networks. First,
to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to
habitat selection and space use (REF)<!-- TODO-->, individual fitness and
phenotypes (Webber & Vander Wal 2018), and movement ecology and collective
movement (Jolles et al. 2019).

@@

@@

## Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974)<!-- TODO-->. As a result, caribou tend to occupy
and re-use craters once they are established and there is considerably less
access to forage than when the landscape is snow-free. Importantly, craters
exist on the landscape at multiple scales: the crater scale (a single crater),
the feeding area scale (multiple craters in close proximity), and the winter
range scale (all craters within an individual’s range) [@Mayor_2009]. The
distribution of forage for caribou is therefore seasonally heterogeneous, with
greater access in snow-free seasons compared to winter as well as spatially
heterogeneous in winter because the distribution of craters on the landscape
varies meaning that access to vegetation in winter is highly variable for
caribou. We therefore partitioned data into two discrete seasons:  summer (3
August - 20 September) and winter (1 January – 17 February) based on caribou
socioecology [@Peignier_2019].

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010)<!-- TODO-->. We did not collar all female caribou in the herd;
however, we assumed that our sample of collared animals was random.

## Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division.
Available landcover classification included nine habitat types at <!-- TODO:
double check --> 10m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat. Foraging habitat
<!-- or just lichen like figs?--> consisted of lichen habitat, open habitat
consisted of wetland, rocky barrens, and anthropogenic habitat types, while
forest habitat consisted of coniferous forest, conifer scrub, broadleaf forest,
and mixed-wood forest habitat types. Water habitat was excluded from all
subsequent analyses.
<!-- using the raster package for R? -->

@@

### Network types and construction (defining layers for the multilayer network)
<!-- should use something like "using syntax from Kivela 2014..." or 
Following [@Kivela_2014]:
other info to maybe include: the multilayer network was not node-aligned since
not all individuals were on all layers
the temporal aspect is ordinal coupling and landcover aspect is categorical coupling

@@

We weighted edges of social networks by the strength of association between
caribou using the simple ratio index (SRI, Cairns and Schwager 1987): <!-- TODO-->

>>>>

# Abstract
Scale remains a seminal concept in ecology. Spatial scale, for instance, has
become a central consideration in the way we understand landscape ecology and
animal space use. Meanwhile, social processes can also scale from fine-scale
interactions to co-occurrence to overlapping home ranges. Multilayer networks
promise the integration of monolayer animal social networks with the complexity
and importance of animal space use and movement in heterogeneous landscapes.
Despite the complex interplay between social networks and how animals use space,
there remains an important biological and methodological gap in our
understanding of an animal’s perception of scale, e.g., grain and extent. But
scale can affect multilayer network dynamics and how we build and interpret
them. Here we discuss the role of scale in the context of multilayer networks
and provide a case study of caribou (*Rangifer tarandus*) to illustrate the role
of social, spatial, and temporal scale on multilayer processes. We also present
perspectives on future development and application of scale in multilayer
networks with respect to habitat selection and space use, individual fitness and
phenotypes, and movement ecology and collective movement. Effective integration
of social and spatial processes, including biologically meaningful scales,
within the context of animal social networks is an emerging area of research;
our contribution uniquely incorporates perspectives on how the social
environment and spatial processes are linked across scales in a multilayer
framework. Based on social network and behavioural ecology theory as well as our
case study, we demonstrate potential for interconnectedness among individual
animals, their social environment, and the landscapes they occupy.

@@

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address two
principal questions associated with scale-dependent multilayer networks. First,
to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to
habitat selection and space use <!-- TODO: (REF)-->, individual fitness and
phenotypes (Webber & Vander Wal 2018), and movement ecology and collective
movement (Jolles et al. 2019).

@@

<!-- TODO: units of spatial scale? -->

@@

## Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974)<!-- TODO: get Bergerud ref-->. As a result, caribou tend to occupy
and re-use craters once they are established and there is considerably less
access to forage than when the landscape is snow-free. Importantly, craters
exist on the landscape at multiple scales: the crater scale (a single crater),
the feeding area scale (multiple craters in close proximity), and the winter
range scale (all craters within an individual’s range) [@Mayor_2009]. The
distribution of forage for caribou is therefore seasonally heterogeneous, with
greater access in snow-free seasons compared to winter as well as spatially
heterogeneous in winter because the distribution of craters on the landscape
varies meaning that access to vegetation in winter is highly variable for
caribou. We therefore partitioned data into two discrete seasons:  summer (3
August - 20 September) and winter (1 January – 17 February) based on caribou
socioecology [@Peignier_2019].

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010)<!-- TODO: get Bjorneraas ref-->. We did not collar all female caribou in the herd;
however, we assumed that our sample of collared animals was random.

## Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division.
Available landcover classification included nine habitat types at <!-- TODO:
double check lc res--> 10m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat. Foraging habitat
<!-- TODO: "foraging" should be "lichen" ?--> consisted of lichen habitat, open habitat
consisted of wetland, rocky barrens, and anthropogenic habitat types, while
forest habitat consisted of coniferous forest, conifer scrub, broadleaf forest,
and mixed-wood forest habitat types. Water habitat was excluded from all
subsequent analyses.
<!-- TODO: cite the raster package for R? -->

@@

### Network types and construction (defining layers for the multilayer network)
<!-- TODO: should use something like "using syntax from Kivela 2014..." or 
Following [@Kivela_2014]:
other info to maybe include: the multilayer network was not node-aligned since
not all individuals were on all layers
the temporal aspect is ordinal coupling and landcover aspect is categorical coupling

@@

We weighted edges of social networks by the strength of association between
caribou using the simple ratio index (SRI, Cairns and Schwager 1987): <!-- TODO: get Cairns ref-->
